Ab Initio Effective One-Electron Potential Operators: Applications for Charge-Transfer Energy in Effective Fragment Potentials

### 2019-11-20 papers/2_oep_technique/paper-r2.tex
Added multiple figures and new tables

\section{\label{ss.5.2}Introduction}

Charge transfer (CT) between molecules occurs when the net electronic populations 
of interacting molecules change which leads to an additional stabilization 
of a molecular aggregate. The effect of CT on the interaction energy 
is known to be often non\hyp{}negligible, especially
in donor\hyp{}acceptor systems such as H\hyp{}bonded species
and charged complexes.
\emph{Ab initio} fragmentation methods, such as
the second generation of the effective fragment potential method (EFP2),
are often used to describe the homogeneous extended molecular systems 
such as solutions
and recently even biomolecules. EFP2 relies on the intermolecular 
perturbation theory, and the total interaction energy is approximated as
a sum of five contributions due to the Coulombic interaction of the unperturbed (isolated)
charge densities of the monomers, exchange\hyp{}repulsion energy,
induction energy due to mutual polarization, intermolecular dispersion energy,
as well as the CT energy,
%
\begin{equation}
 U^{\rm EFP2} \approx U^{\rm Coul} + U^{\rm Ex-Rep} + U^{\rm Ind} + U^{\rm Disp} + U^{\rm CT} \;.
\end{equation}
%
However, calculating the CT stabilization energy is usually much more expensive 
than evaluating the contributions due to other effects.\cite{Gordon.Fedorov.Pruitt.Slipchenko.ChemRev.2012}
%However, the computational cost of the CT term is usually at least few times larger
%than the next costly contribution $U^{\rm Ex-Rep}$, which is still too costly
%to perform molecular dynamics (MD) simulations including the $U^{\rm CT}$. 
In effect,
this part of the EFP2 energy is usually ignored in most of the applications. In the 
recent LIBEFP library, CT term were even not implemented due to those reasons.

In this work, we propose an alternative formulation of the CT energy compatible with the
EFP2 method. We start from the perturbation theory of Murrell et al.
and
introduce the effective one\hyp{}electron potential (OEP) operators
to eliminate the electron\hyp{}repulsion integrals (ERI's) from the original theory.
Later we validate the model against the EFP2 method in Section XXX. Finally,
we conclude this work and provide few remarks on future development in Section XXX.

\section{\label{s:2}Effective One-Electron Potential Operators}

To establish the notation within the present work, 
the formalism of one\hyp{}electron potential (OEP) operators is
given first.
The one\hyp{}electron Coulomb static effective potential $v^{\rm eff}({\bf r})$
produced by a certain effective one\hyp{}electron charge density distribution $\rho^{\rm eff}({\bf r})$
is 
%
\begin{equation} \label{e:v-eff}
	v^{\rm eff}({\bf r}) = \int \frac{ \rho^{\rm eff}({\bf r}') }{ \vert {\bf r}' - {\bf r} \vert} d{\bf r}' \;,
\end{equation}
%
where ${\bf r}$ is a spatial coordinate. 
For convenience, the total effective density can be split into nuclear and electronic contributions,
%
\begin{equation} \label{e:rho-eff}
 \rho^{\rm eff}({\bf r}) = \lambda \rho^{\rm eff}_{\rm nuc}({\bf r}) + \rho^{\rm eff}_{\rm el}({\bf r}) \;,
\end{equation}
%
where $\lambda$ is a certain parameter and is assumed either 1 or 0 in this work.
Formally, the electronic part can be expanded in terms of an effective
one\hyp{}particle density matrix represented in a certain basis of orbitals, $\BM{\phi}({\bf r})$,
%
\begin{equation} \label{e:d-spectral}
	\rho^{\rm eff}_{\rm el}({\bf r}) = \sum_{\alpha\beta} D_{\alpha\beta}^{\rm eff} 
	\phi_\alpha({\bf r}) \phi_\beta^{*}({\bf r})  \;.
\end{equation}
%
Based on that, the operator form of the effective potential 
can be written as
%
\begin{equation} \label{e:oep-operator}
	\hat{v}^{\rm eff} = 
        \lambda \hat{v}_{\rm nuc} +
        %\sum_{\alpha\beta} 
	%\Ket{\alpha} 
        \int d{\bf r} \Ket{{\bf r}} 
        v^{\rm eff}({\bf r})
	%\left[
	%\int
	%\hat{J}^{\rm}_{\alpha\beta}({\bf r}') 
	%\rho^{\rm eff}({\bf r}')
	%d{\bf r}' 
	%\right]
	%\Bra{\beta}  
        \Bra{{\bf r}}
\end{equation}
%
with the nuclear
%and Coulomb operators 
operator
defined by
%
%\begin{subequations}
%\begin{align}
\begin{equation}
        \hat{v}_{\rm nuc} \equiv \sum_x^{\rm At}  
                     \int d{\bf r} \Ket{{\bf r}} 
                     \frac{Z_x}{\vert {\bf r} - {\bf r}_x \vert}
                     \Bra{{\bf r}} \;.
	%\hat{J}^{\rm}_{\alpha\beta}({\bf r}) f({\bf r}) &\equiv
	%f({\bf r}) 
	%\int
	%\frac{ \phi_\alpha^{*}({\bf r}') \phi_\beta({\bf r}') }{ \vert {\bf r}' - {\bf r} \vert} d{\bf r}'
\end{equation}
%\end{align}
%\end{subequations}
%
%for any one\hyp{}electron function $f({\bf r})$. 
and $\BraKet{\bf r}{\alpha} = \varphi_\alpha({\bf r})$.
The matrix element of the effective potential operator
is therefore given by
%
\begin{equation}
	\tBraKet{\alpha}{ \hat{v}^{\rm eff} }{\beta}
	= \lambda \sum_x^{\rm At} W_{\alpha\beta}^{(x)} +
        \sum_{\gamma\delta} \BraKet{\alpha\beta}{\gamma\delta} D^{\rm eff}_{\gamma\delta}  \;,
\end{equation}
%
where 
%
\begin{equation}
 W_{\alpha\beta}^{(x)} = 
 Z_x \int \frac{\phi^*_\alpha({\bf r}) \phi_\beta({\bf r})}{\vert {\bf r} - {\bf r}_x \vert} d{\bf r} \;,
\end{equation}
%
$Z_x$ is the atomic number of the $x$th atom,
and the symbol `At' denotes all atoms that contribute to the effective potential.
In the above equation and throughout the work, 
%
\begin{equation}
\tBraKet{\alpha}{\mathscr{O}(1)}{\beta} \equiv \int \phi^*_\alpha({\bf r}) \mathscr{O}(1) \phi_\beta({\bf r}) d{\bf r} 
\end{equation}
%
for any one\hyp{}electron operator $\mathscr{O}(1)$ 
and the electron repulsion integral (ERI)
is defined according to
%
\begin{equation}
	\BraKet{\alpha\beta}{\gamma\delta} \equiv
	\iint 
	\frac{ \phi_\alpha^{*}({\bf r}_1) \phi_\beta({\bf r}_1) 
	       \phi_\gamma^{*}({\bf r}_2) \phi_\delta({\bf r}_2) }{ \vert {\bf r}_1 - {\bf r}_2 \vert}
	d{\bf r}_1 d{\bf r}_2  \;.
\end{equation}
%

%\subsection{\label{s.334}Practical OEP-Based Calculations}

\subsection{\label{ss:2.3}Incorporating Electron Repulsion Integrals into Effective Potentials}

Consider now an arbirtary functional $\mathcal{F}$ that explicitly depends on the 
ERI's. In this work, OEP's are defined by the following transformation
%
 \begin{equation}
 \mathcal{F}
 \left[ 
   \BraKet{ij}{k^Al^A}
	 \right] = \tBraKet{i}{\hat{v}_{kl}^A}{j}  \;,
 \end{equation}
%
where 
% $A$ and $B$ denote different molecules and $\phi_i$ is the $i$th molecular orbital
%or basis function.
$\hat{v}_{kl}^A$ is the effective one-electron potential operator given by Eq.~\eqref{e:oep-operator} 
with the
effective density $\rho_{kl}^A({\bf r}) \equiv \phi_k^A({\bf r})\phi_l^A({\bf r})$.
The summations over $k$ and $l$ can be incorporated into the total effective one-electron potential operator
$\hat{v}_{\text{eff}}^A$
to produce
%
\begin{equation}
	\sum_{ij}\sum_{kl\in A} \mathcal{F}\left[ 
   \BraKet{ij}{k^Al^A}
 \right] = \sum_{ij} \tBraKet{i}{\hat{v}_{\text{eff}}^A}{j}  \;.
\end{equation}
%
Thus, the total computational effort is, in principle, reduced from the fourth-fold
sum involving evaluation of ERI's to the two-fold sums of cheaper one-electron integrals.
It is also possible to generalize the above expression even further by
summing over all possible functionals ${\mathcal{F}}_t$
%
\begin{equation} \label{e:ft-reduction}
	\sum_t \sum_{ij}\sum_{kl\in A} {\mathcal{F}}_t\left[ 
   \BraKet{ij}{k^Al^A}
 \right] = \sum_{ij} \tBraKet{i}{\hat{v}_{\text{eff}}^A}{j} \;.
\end{equation}
%
The above design has the advantage that it opens the possibility to define first\hyp{}principles
effective fragments as long as the $P_{ij}$ and $P_{ijkl}$ 
from Eq.~\eqref{e:exp-val-series} are computable and can be approximately
partitioned in between the interacting fragments.
%the resulting effective potentials are fully first\hyp{}principles
%and no extensive case\hyp{}dependent fitting procedures are necessary as long as the $P_{ij}$ and $P_{ijkl}$
%are computable.
%Only the effective potentials of \emph{independent} fragment $A$ need to be determined once and for all
%and stored in a file. 
%Note also, that, in principle, there is no approximation 
%made here at that moment.

@@

\subsection{\label{s.333}Distributed Multipole Expansion of OEP's}

This scheme is most applicable in the case of matrix elements of the type
$
 \tBraKet{j^B}{\hat{v}_{\rm eff}^A}{l^B}
$
because it can be considered as a Coulombic interaction between $\rho^B_{jl}$
and $\rho^A_{\rm eff}$. 
In general, given certain two effective one\hyp{}electron density distributions,
the associated effective Coulombic interaction energy can be estimated from the classical formula
according to
%
\begin{equation}
 \Delta^{(XY)}_{\rm eff} E_{\rm Coul} = \iint \frac{\rho_{\rm eff}^X({\bf r}_1) \rho_{\rm eff}^Y({\bf r}_2)}
 {\vert {\bf r}_1 - {\bf r}_2 \vert} 
d{\bf r}_1 d{\bf r}_2 \;.
\end{equation}
%
The above integral can be approximated by the
distributed multipole (DMTP)
expansions of the respective densities as follows
%
\begin{multline}
  \Delta^{(XY)}_{\rm eff} E_{\rm Coul} \approx
 \rho_{\rm eff}^X \odot \rho_{\rm eff}^Y \equiv
 \sum_{u\in A} \sum_{w\in B} \Big\{ 
 T^{(0)}_{uw}
 q_{\rm eff}^{(u)}  q_{\rm eff}^{(w)} \\
 - {\bf T}^{(1)}_{uw} \cdot 
   \left[ q_{\rm eff}^{(u)} {\BM\upmu}_{\rm eff}^{(w)} - q_{\rm eff}^{(w)} {\BM\upmu}_{\rm eff}^{(u)} \right]
 - {\bf T}^{(2)}_{uw} : 
  {\BM\upmu}_{\rm eff}^{(u)}  \otimes {\BM\upmu}_{\rm eff}^{(w)} 
 \ldots
 \Big\}
\end{multline}
%
In the above equations, $q_{\rm eff}^{(u)}$ and ${\BM\upmu}^{(u)}_{\rm eff}$ 
are the effective distributed monopole (charge)
and dipole moment, respectively, centered on the $u$th site, 
whereas ${{\bf T}^{(r)}_{uw}}$ are the so called interaction tensors or rank $r$ 
that can be found elsewhere.
The symbol `$\odot$' denotes the sum of all the tensor contractions
performed over the DMTP's of molecule $X$ and $Y$ to yield the associated interaction energy.
The choice of the distribution centres as well as the truncation order of the multipole expansion
is crucial in compromising the accuracy and computational cost of the resulting expressions.
There is many ways in which this can be achieved, e.g., through the distributed multipole analysis (DMA)
of Stone and Anderton, %cite
the cumulative atomic multipole moments of Sokalski and Poirier, %cite
the localised distributed multipole expansion (LMTP) of Etchtebest et al. %cite
or schemes based on fitting to electrostatic potential.

\subsection{\label{s.3344}Density Fitting of OEP's}

@@

\subsubsection{Density Fitting in Nearly-Complete Space}

>>>>

\section{\label{s:1.introduction}Introduction}

Charge transfer (CT) between molecules occurs when the net electronic populations 
of interacting molecules change which leads to an additional stabilization 
of a molecular aggregate.\cite{Otto.Ladik.IJQC.1980,Stone.TheTheoryOfIntermolecularForces.1996}
Although CT processes can be relatively easily scrutinized based
on the total amount of charge transferred between interacting species,\cite{Otto.Ladik.IJQC.1980}
evaluation of its contribution to the intermolecular interaction
potential\cite{Stone.TheTheoryOfIntermolecularForces.1996} is far from trivial due to 
its very complex quantum mechanical (QM) origins even at the Hartree\hyp{}Fock approximation level,\cite{Otto.Ladik.ChemPhys.1975}
and cannot be realized in terms of any classical nor semi\hyp{}classical approach.
In fact, CT does not naturally emerge in the symmetry 
adapted perturbation theory\cite{Jeziorski.Moszynski.Szalewicz.ChemRev.1994} (SAPT)
due to certain technical aspects related with the use of finite basis set expansions
in modern quantum chemistry calculations and minimizing the basis set superimposition error (BSSE). 
Recently, Stone and Misquitta showed that CT energy
can be extracted from SAPT calculations by comparing the induction energies
of fictional systems in which basis functions are centered either on the monomers only, or the entire 
interacting complex.\cite{Stone.Misquitta.CPL.2009}
CT energy was formulated by Murrell et al.~\cite{Murrell.Randic.Williams.Longuet-Higgins.ProcRSocLondA.1965} 
in their perturbation theory in the region of small wavefunction overlap up to second order.
However, their theory is also computationally too expensive to apply for efficient
calculation of intermolecular forces in molecular dynamics because
it involces instantaneous calculations of the electron repulsion integrals (ERI's)
in molecular orbital (MO) basis.\cite{Otto.Ladik.ChemPhys.1975}

The apparent difficulty in theoretically characterising the CT energy
in terms of the interacting molecular fragments
is indeed a challenge in developing modern force fields 
or \emph{ab initio} fragmentation methods\cite{Gordon.Fedorov.Pruitt.Slipchenko.ChemRev.2012}
for modeling structure
and dynamics in condensed phases.
This needs to be contrasted with the
Coulombic electrostatics,
non\hyp{}Coulombic repulsion (i.e., due to Pauli exclusion principle), 
dispersion and induction for instance, 
which to a certain extent
can be well described by, respectively, 
the multipole moments of charge density distributions,\cite{Sokalski.Poirier.CPL.1983,
Etchebest.Lavery.Pullman.TheorChimActa.1982,Stone.JCTC.2005}
the van der Waals repulsive and attractive potentials, 
as well as the polarizability models, i.e., all being relatively simple and computationally
inexpensive to evaluate mathematical models. 
Due to this reason, the
CT effects 
%while usually of significant and non\hyp{}negligible magnitude as compared 
%to the total interaction energy, 
are not included in any molecular mechanics force field developed up to date
such as OPLS, Amber, CHARMM, Gromos or even polarizale AMOEBA force field. 
The only way to
incorporate the CT effects in the condensed\hyp{}phase simulations,
apart from performing full QM molecular dynamics simulations,
is through the second generation of the effective fragment potential 
(EFP2) method,\cite{Gordon.Smith.Xu.Slipchenko.AnnuRevPhysChem.2013,
   Nguyen.Pachter.Day.JCP.2014,
   Day.Jensen.Gordon.Webb.Stevens.Krauss.Garmer.Basch.Cohen.JCP.1996}
developed by Gordon group.\cite{Li.Gordon.Jensen.JCP.2006,Xu.Gordon.JCP.2013}
EFP2 is an \emph{ab initio} force field
derived on the grounds of the first\hyp{}principles perturbation theory
at the HF level and including the intermolecular dispersion effects
by the response theory based on the imaginary frequency\hyp{}dependent dipole\hyp{}dipole polarizabilities.
That is to say, the total intermolecular 
interaction potential is approximated as 
%
\begin{equation}\label{e:u-efp2}
 U^{\rm EFP2} \approx U^{\rm Coul} + U^{\rm Ex-Rep} + U^{\rm Ind} + U^{\rm Disp} + U^{\rm CT} \;.
\end{equation}
%
where $U^{\rm Coul}$ is the Coulombic interaction energy of the unperturbed charge\hyp{}density
distributions of the monomers, approximated by the distributed multipole approximation,
$U^{\rm Ex-Rep}$ is the exchange\hyp{}repulsion energy originating from the Pauli exclusion
principle, that could be roughly understood as the Heitler\hyp{}London interaction
energy with $U^{\rm Coul}$ subtracted, $U^{\rm Ind}$ and $U^{\rm Disp}$
are the induction and dispersion energies obtained from the distributed polarizability
approximation, and finally $U^{\rm CT}$ is the CT energy, the focus of this work.

Despite the considerable success of the EFP2 theory
in accurately modeling the extended molecular systems 
like solutions and recently even biomolecules
with the level of accuracy
roughly arguably comparable to the M{\o}ller\hyp{}Plesset perturbation theory,
evaluation of the CT energy in EFP2 model is still too costly for
typical uses in the
molecular dynamics simulations. It has been reported that 
the implementation of the EFP2 CT energy component in GAMESS
is on average 3-10 times more demanding than the exchange\hyp{}repulsion
component, becoming a bottleneck of evaluation of total interaction 
energy and gradients.\cite{Gordon.Smith.Xu.Slipchenko.AnnuRevPhysChem.2013,
Li.Gordon.Jensen.JCP.2006}
In effect, the CT energy component is usually ignored in most of the applications. In the 
recent LIBEFP library, CT term were not implemented in the official release 
of the code.\cite{Kaliman.Slipchenko.JCC.2015}

%However, calculating the CT stabilization energy is usually much more expensive 
%than evaluating the contributions due to other effects.
%However, the computational cost of the CT term is usually at least few times larger
%than the next costly contribution $U^{\rm Ex-Rep}$, which is still too costly
%to perform molecular dynamics (MD) simulations including the $U^{\rm CT}$.

The effect of CT on the interaction energy 
is known to be often non\hyp{}negligible, especially
in donor\hyp{}acceptor systems such as H\hyp{}bonded species
and charged complexes, although highly depends on the theoretical
approach being used. 
The approach adopted by the founders of EFP2 model
is based on
%The most common methodology,
%adopted also by the founders of EFP2 model,
%is 
the approximation to SAPT at HF level of theory 
which is based on the isolated monomeric wave functons
as references (here abbreviated to CT/HF0).
In this formulation, the CT energy can be understood
as a part of the polarization energy that cannot be
linked to dispersion and pure induction interactions. 
In this definition of CT, mixing of virtual and occupied MO's
technically belonging to different unperturbed wavefunction occurs,
unlike the pure induction and dispersion, in which such mixings are contained
solely within one monomer.
%This level of definition of the CT energy will be referred here as to
%the IPTO-CT energy. 
%For defining the CT/HF0, 
%most common reference is the HF approximation
%most certainly because the HF energy functional is known
%and its properties are well understood.
%In the EFP2 model, the total interaction energy is approximated as
%a sum of five contributions due to the Coulombic interaction of the unperturbed (isolated)
%charge densities of the monomers, exchange\hyp{}repulsion energy,
%induction energy due to mutual polarization, intermolecular dispersion energy,
%as well as the CT energy.
%For developing the efficient 
%\emph{ab initio} fragmentation method 
%
% the second generation of the effective fragment potential method (EFP2),
%the Hartree\hyp{}Fock (HF) model or the density functional theory (DFT) were
%employed as starting points 
%Modelling the structure and dynamics of extended molecular aggregates, such as
%liquids and solutions require
%EFP2 relies on the intermolecular 
%perturbation theory, and the total interaction energy is approximated as ...
%
%\begin{equation}
% U^{\rm EFP2} \approx U^{\rm Coul} + U^{\rm Ex-Rep} + U^{\rm Ind} + U^{\rm Disp} + U^{\rm CT} \;.
%\end{equation}
%%

One of the main goals of this work, apart from developing
more efficient model of the CT/HF0 energy, is to
extend the effective one\hyp{}electron potentials (here referred to as the OEP) 
technique
used previously\cite{
Otto.Ladik.ChemPhys.1975,
Holas.March.PhysRevA.1991,
Weber.Thiel.TCA.2000,
Neese.JCP.2005,
Cisneros.Andres.Piquemal.Darden.JCP.2005,
Piquemal.Cisneros.Reinhardt.Gresh.Darden.JCP.2006,
Li.Gordon.Jensen.JCP.2006,
Blasiak.Lee.Cho.JCP.2013,
Blasiak.Maj.Cho.Gora.JCTC.2015}
to simplify the rigorous and costly quantum chemistry models of extended systems
with a particular emphasis on solvation phenomena and molecular dynamics.
The presented OEP technique of removing ERI's from the working
equations
follows the notion of the importance of 
one\hyp{}electron densities in chemistry\cite{Holas.March.PhysRevA.1991},
thus reducing the complicated summations involving ERI's to much shorter expressions involving
only one\hyp{}electron integrals.
Therefore, the OEP computational method is first presented in Section~II.
Next, in Section III,
the new technique is used to derive an alternative formulation of the CT/HF0 energy
compatible with the EFP2 method. Subsequently, after details of computations
are discussed in Section IV, in Section V the validation of the 
OEP\hyp{}based CT/HF0 model is presented and its performance in terms 
of accuracy and computational speed is compared against the EFP2
model. Finally, few concluding remarks and comments for future work
are given in Section VI.
%ERI's, which are oftentimes the main bottleneck
%in simulating large systems, will be projected out from the working equations
%to ultimately form the effective fragments in a form of small matrices that are much easier to handle.
%The proposed methodology aims in removing ERI's from the working equations
%to ultimately form the effective fragments that are much easier to handle.
%
%It is believed that the proposed concept can be helpful in designing future methods tuned 
%for studying electronically excited states in solutions, better describing molecular
%dynamics as well as enabling the vibrational analysis of extended systems
%In this work, we propose an alternative formulation of the CT energy compatible with the
%EFP2 method. We start from the perturbation theory of Murrell et al.
%and
%introduce the effective one\hyp{}electron potential (OEP) operators
%to eliminate the electron\hyp{}repulsion integrals (ERI's) from the original theory.
%Later we validate the model against the EFP2 method in Section XXX. Finally,
%we conclude this work and provide few remarks on future development in Section XXX.

\section{\label{s:2.oep}Effective One-Electron Potential Operators}

To establish the notation within the present work, 
the formalism of one\hyp{}electron potential (OEP) operators is
given first.
The one\hyp{}electron Coulomb static effective potential $v^{\rm eff}({\bf r})$
produced by a certain effective one\hyp{}electron charge density distribution $\rho^{\rm eff}({\bf r})$
is 
%
\begin{equation} \label{e:v-eff}
	v^{\rm eff}({\bf r}) = \int \frac{ \rho^{\rm eff}({\bf r}') }{ \vert {\bf r}' - {\bf r} \vert} d{\bf r}' \;,
\end{equation}
%
where ${\bf r}$ is a spatial coordinate. 
For convenience, the total effective density can be split into nuclear and electronic contributions,
%
\begin{equation} \label{e:rho-eff}
 \rho^{\rm eff}({\bf r}) = \lambda \rho^{\rm eff}_{\rm nuc}({\bf r}) + \rho^{\rm eff}_{\rm el}({\bf r}) \;,
\end{equation}
%
where $\lambda$ is a certain parameter and is assumed either 1 or 0 in this work.
Formally, the electronic part can be expanded in terms of an effective
bond order matrix represented in a certain basis of orbitals, $\BM{\phi}({\bf r})$,
%
\begin{equation} \label{e:d-spectral}
	\rho^{\rm eff}_{\rm el}({\bf r}) = \sum_{\alpha\beta} P_{\alpha\beta}^{\rm eff} 
	\phi_\alpha({\bf r}) \phi_\beta^{*}({\bf r})  \;.
\end{equation}
%
Based on that, the operator form of the effective potential 
can be written as
%
\begin{equation} \label{e:oep-operator}
	\hat{v}^{\rm eff} = 
        \lambda \hat{v}_{\rm nuc} +
        %\sum_{\alpha\beta} 
	%\Ket{\alpha} 
        \int d{\bf r} \Ket{{\bf r}} 
        v^{\rm eff}({\bf r})
	%\left[
	%\int
	%\hat{J}^{\rm}_{\alpha\beta}({\bf r}') 
	%\rho^{\rm eff}({\bf r}')
	%d{\bf r}' 
	%\right]
	%\Bra{\beta}  
        \Bra{{\bf r}}
\end{equation}
%
with the nuclear
%and Coulomb operators 
operator
defined by
%
%\begin{subequations}
%\begin{align}
\begin{equation} \label{e:oep-operator-nuc}
        \hat{v}_{\rm nuc} \equiv \sum_x^{\rm At}  
                     \int d{\bf r} \Ket{{\bf r}} 
                     \frac{Z_x}{\vert {\bf r} - {\bf r}_x \vert}
                     \Bra{{\bf r}} \;.
	%\hat{J}^{\rm}_{\alpha\beta}({\bf r}) f({\bf r}) &\equiv
	%f({\bf r}) 
	%\int
	%\frac{ \phi_\alpha^{*}({\bf r}') \phi_\beta({\bf r}') }{ \vert {\bf r}' - {\bf r} \vert} d{\bf r}'
\end{equation}
%\end{align}
%\end{subequations}
%
%for any one\hyp{}electron function $f({\bf r})$. 
and $\BraKet{\bf r}{\alpha} = \varphi_\alpha({\bf r})$.
The matrix element of the effective potential operator
is therefore given by
%
\begin{equation} \label{e:oep-matrix-element}
	\tBraKet{\alpha}{ \hat{v}^{\rm eff} }{\beta}
	= \lambda \sum_x^{\rm At} W_{\alpha\beta}^{(x)} +
        \sum_{\gamma\delta} \BraKet{\alpha\beta}{\gamma\delta} P^{\rm eff}_{\gamma\delta}  \;,
\end{equation}
%
where 
%
\begin{equation} \label{e:oep-matrix-element-nuc}
 W_{\alpha\beta}^{(x)} = 
 Z_x \int \frac{\phi^*_\alpha({\bf r}) \phi_\beta({\bf r})}{\vert {\bf r} - {\bf r}_x \vert} d{\bf r} \;,
\end{equation}
%
$Z_x$ is the atomic number of the $x$th atom,
and the symbol `At' denotes all atoms that contribute to the effective potential.
In the above equation and throughout the work, 
%
\begin{equation} \label{e:1el-operator}
\tBraKet{\alpha}{\mathscr{O}(1)}{\beta} \equiv \int \phi^*_\alpha({\bf r}) \mathscr{O}(1) \phi_\beta({\bf r}) d{\bf r} 
\end{equation}
%
for any one\hyp{}electron operator $\mathscr{O}(1)$ 
and the electron repulsion integral (ERI)
is defined according to
%
\begin{equation} \label{e:eri}
	\BraKet{\alpha\beta}{\gamma\delta} \equiv
	\iint 
	\frac{ \phi_\alpha^{*}({\bf r}_1) \phi_\beta({\bf r}_1) 
	       \phi_\gamma^{*}({\bf r}_2) \phi_\delta({\bf r}_2) }{ \vert {\bf r}_1 - {\bf r}_2 \vert}
	d{\bf r}_1 d{\bf r}_2  \;.
\end{equation}
%

\subsection{\label{ss:2.1.oep-technique}Incorporating Electron Repulsion Integrals into Effective Potentials}

Consider now an arbirtary functional $\mathcal{F}$ that explicitly depends on the 
ERI's. In this work, OEP's are defined by the following transformation
%
 \begin{equation}
 \mathcal{F}
 \left[ 
   \BraKet{ij}{k^Al^A}
	 \right] = \tBraKet{i}{\hat{v}_{kl}^A}{j}  \;,
 \end{equation}
%
where 
% $A$ and $B$ denote different molecules and $\phi_i$ is the $i$th molecular orbital
%or basis function.
$\hat{v}_{kl}^A$ is the effective one-electron potential operator given by Eq.~\eqref{e:oep-operator} 
with the
effective density $\rho_{kl}^A({\bf r}) \equiv \phi_k^A({\bf r})\phi_l^A({\bf r})$.
The summations over $k$ and $l$ can be incorporated into the total effective one-electron potential operator
$\hat{v}_{\text{eff}}^A$
to produce
%
\begin{equation}
	\sum_{ij}\sum_{kl\in A} \mathcal{F}\left[ 
   \BraKet{ij}{k^Al^A}
 \right] = \sum_{ij} \tBraKet{i}{\hat{v}_{\text{eff}}^A}{j}  \;.
\end{equation}
%
Thus, the total computational effort is, in principle, reduced from the fourth-fold
sum involving evaluation of ERI's to the two-fold sums of cheaper one-electron integrals.
It is also possible to generalize the above expression even further by
summing over all possible functionals ${\mathcal{F}}_t$
%
\begin{equation} \label{e:ft-reduction}
	\sum_t \sum_{ij}\sum_{kl\in A} {\mathcal{F}}_t\left[ 
   \BraKet{ij}{k^Al^A}
 \right] = \sum_{ij} \tBraKet{i}{\hat{v}_{\text{eff}}^A}{j} \;.
\end{equation}
%
The above design has the advantage that it opens the possibility to define first\hyp{}principles
effective fragments as long as 
the functionals ${\mathcal{F}}_t$ 
%the $P_{ij}$ and $P_{ijkl}$ 
%from Eq.~\eqref{e:exp-val-series} 
are computable and can be approximately
partitioned in between the interacting fragments.
%the resulting effective potentials are fully first\hyp{}principles
%and no extensive case\hyp{}dependent fitting procedures are necessary as long as the $P_{ij}$ and $P_{ijkl}$
%are computable.
%Only the effective potentials of \emph{independent} fragment $A$ need to be determined once and for all
%and stored in a file. 
%Note also, that, in principle, there is no approximation 
%made here at that moment.

@@

\subsection{\label{ss:2.2.oep-DMTP}Distributed Multipole Expansion of OEP's}

This scheme is most applicable in the case of matrix elements of the type
$
 \tBraKet{j^B}{\hat{v}_{\rm eff}^A}{l^B}
$
because it can be considered as a Coulombic interaction between $\rho^B_{jl}$
and $\rho^A_{\rm eff}$. 
In general, given certain two effective one\hyp{}electron density distributions,
the associated effective Coulombic interaction energy can be estimated from the classical formula
according to
%
\begin{equation} \label{e:coulomb-energy:general}
 \Delta^{(XY)}_{\rm eff} E_{\rm Coul} = \iint \frac{\rho_{\rm eff}^X({\bf r}_1) \rho_{\rm eff}^Y({\bf r}_2)}
 {\vert {\bf r}_1 - {\bf r}_2 \vert} 
d{\bf r}_1 d{\bf r}_2 \;.
\end{equation}
%
The above integral can be approximated by the
distributed multipole (DMTP)
expansions of the respective densities as follows
%
\begin{multline} \label{e:coulomb-energy:dmtp}
  \Delta^{(XY)}_{\rm eff} E_{\rm Coul} \approx
 \rho_{\rm eff}^X \odot \rho_{\rm eff}^Y \equiv
 \sum_{u\in A} \sum_{w\in B} \Big\{ 
 T^{(0)}_{uw}
 q_{\rm eff}^{(u)}  q_{\rm eff}^{(w)} \\
 - {\bf T}^{(1)}_{uw} \cdot 
   \left[ q_{\rm eff}^{(u)} {\BM\upmu}_{\rm eff}^{(w)} - q_{\rm eff}^{(w)} {\BM\upmu}_{\rm eff}^{(u)} \right]
 - {\bf T}^{(2)}_{uw} : 
  {\BM\upmu}_{\rm eff}^{(u)}  \otimes {\BM\upmu}_{\rm eff}^{(w)} 
 \ldots
 \Big\}
\end{multline}
%
In the above equations, $q_{\rm eff}^{(u)}$ and ${\BM\upmu}^{(u)}_{\rm eff}$ 
are the effective distributed monopole (charge)
and dipole moment, respectively, centered on the $u$th site, 
whereas ${{\bf T}^{(r)}_{uw}}$ are the so called interaction tensors or rank $r$ 
that can be found elsewhere.
The symbol `$\odot$' denotes the sum of all the tensor contractions
performed over the DMTP's of molecule $X$ and $Y$ to yield the associated interaction energy.
The choice of the distribution centres as well as the truncation order of the multipole expansion
is crucial in compromising the accuracy and computational cost of the resulting expressions.
There is many ways in which this can be achieved, e.g., through the distributed multipole analysis (DMA)
of Stone\cite{Stone.JCTC.2005}
the cumulative atomic multipole moments of Sokalski and Poirier\cite{Sokalski.Poirier.CPL.1983}
the localised distributed multipole expansion (LMTP) of Etchtebest, Lavery and Pullman\cite{Etchebest.Lavery.Pullman.TheorChimActa.1982}
or schemes based on fitting to electrostatic potential such as ChelpG\cite{Breneman.Wiberg.JCC.1990} for instance.

\subsection{\label{ss:2.3.oep-GDF}Density Fitting of OEP's}

@@

\subsubsection{\label{sss:2.3.1.GDF-1}Density Fitting in Nearly-Complete Space}
